---
title: "Star Wars data"
author: "Seth Green"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is just some fake data to show where to put things. Just showing some of the demo stuff from [dplyr](https://dplyr.tidyverse.org/#usage) so that we have something to render.

**The idea is to put each analysis question in it's own repo.** This is an analysis of the question: "Who are the characters from Tatooine, Alderaan, and Kashyyyk?"

```{r, message = F}
# first you need to install these packages
library(tidyverse)
```

## Analysis

Not sure how much narrative "report" we want in these docs, and also how much code to include, but both `.Rmd` and `.ipynb` can support whatever format we want, and then render to either an HTML or pdf.

Then, once the analysis is finished and merged to `main`, we can drop that rendered HTML or pdf in `code-for-charlottesville.github.io/LAJC-expungement` and send that link to whoever needs to see it.

```{r cars}
df <- starwars
head(df)
```

```{r}

df %>% 
  filter(homeworld %in% c("Tatooine", "Alderaan", "Kashyyyk")) %>%
  mutate(name, bmi = mass / ((height / 100)  ^ 2)) %>%
  select(name:mass, bmi, species) %>%
  ggplot(aes(x = height, y = log(bmi))) +
    geom_point(aes(colour = species, size = log(mass))) +
    geom_text(aes(label = name), hjust = 1, vjust = 0)
```


```{r}

starwars %>%
  group_by(species) %>%
  summarise(
    n = n(),
    mass = mean(mass, na.rm = TRUE)
  ) %>%
  filter(
    n > 1,
    mass > 50
  ) %>%
  rename(
    `Number of Characters` = n,
    `Average Mass` = mass
    ) %>%
  knitr::kable()

```


