---
title: "Current State of Expungeable Records - Exploring Edge Cases"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r, message=F}
# first you need to install these packages
library(tidyverse)
library(data.table)
library(fs)
library(here)
source(here("code", "helper-functions.R"))
```

There are some edge cases in the data that we want to examine to see just how common they are so we can decide what to do with them.

```{r}
criminal_files <- fs::dir_ls("/home/rstudio/courtdata") %>% str_subset("criminal")
print(paste("Processing", length(criminal_files), "files..."))

res <- imap(criminal_files,  ~ {
  if (.y %% 10 == 0) cat(paste("$", .y, "$"))
  
  .d <- suppressWarnings(read_court_file(.x))
  
  list(
    DispositionCode = table(.d$DispositionCode),
    ChargeType = table(.d$ChargeType),
    Plea = table(.d$Plea),
    CodeNA = sum(.d$CodeSection == "", na.rm = T) + sum(is.na(.d$CodeSection))
  )
})
cat("\n")
```

# Uncommon and Missing Disposition Codes

Looking at the counts of disposition codes across all files.

```{r}
res_dc <- res %>%
  map_dfr(~.x[["DispositionCode"]]) %>% 
  rename(MISSING = `...1`) %>%
  mutate(
    across(everything(), ~replace_na(.x, 0)),
    file = criminal_files
  ) %>%
  select(file, everything()) 

# saving out the raw data
write_csv(res_dc, here("analysis/current-state-of-expungeable-records/data", "counts_DispositionCode.csv"))

res_dc %>%
  select(-file) %>%
  summarise(across(everything(), sum)) %>%
  pivot_longer(everything(), names_to = "DispositionCode", values_to = "Count") %>%
  arrange(desc(Count)) %>%
  mutate(Count = format(Count, big.mark = ",")) %>%
  knitr::kable()
```


# Uncommon Charge Types

Looking at the counts of charge type across all files.

```{r}
res_ct <- res %>%
  map_dfr(~.x[["ChargeType"]]) %>% 
  mutate(
    across(everything(), ~replace_na(.x, 0)),
    file = criminal_files
  ) %>%
  select(file, everything()) 

# saving out the raw data
write_csv(res_ct, here("analysis/current-state-of-expungeable-records/data", "counts_ChargeType.csv"))

res_ct %>%
  select(-file) %>%
  summarise(across(everything(), sum)) %>%
  pivot_longer(everything(), names_to = "ChargeType", values_to = "Count") %>%
  arrange(desc(Count)) %>%
  mutate(Count = format(Count, big.mark = ",")) %>%
  knitr::kable()
```

# Uncommon Pleas

Looking at the counts of hearing plea across all files.

```{r}
res_ct <- res %>%
  map_dfr(~.x[["Plea"]]) %>% 
  mutate(
    across(everything(), ~replace_na(.x, 0)),
    file = criminal_files
  ) %>%
  select(file, everything()) 

# saving out the raw data
write_csv(res_ct, here("analysis/current-state-of-expungeable-records/data", "counts_Plea.csv"))

res_ct %>%
  select(-file) %>%
  summarise(across(everything(), sum)) %>%
  pivot_longer(everything(), names_to = "Plea", values_to = "Count") %>%
  arrange(desc(Count)) %>%
  mutate(Count = format(Count, big.mark = ",")) %>%
  knitr::kable()
```


# Missing Code Section

```{r}
res_cna <- res %>%
  map_int(~.x[["CodeNA"]]) %>%
  sum()

print(paste(res_cna, "missing CodeSection in full data"))
```

# Examples of rare dispositions

Looking for examples of these

* Complied with Law
* Resolved
* Certified Misdemeanor
* Granted
* Denied

```{r, eval=FALSE}
# don't eval when knitting

our_dispositions <- c(
  "Complied With Law",
  "Resolved",
  "Certified Misdemeanor",
  "Granted",
  "Denied"
)

criminal_files <- fs::dir_ls("/home/rstudio/courtdata") %>% str_subset("criminal")
print(paste("Processing", length(criminal_files), "files..."))

res <- imap_dfr(criminal_files,  ~ {
  if (.y %% 10 == 0) cat(paste("$", .y, "$"))
  
  suppressWarnings(read_court_file(.x)) %>%
    filter(DispositionCode %in% our_dispositions)
})
cat("\n")

```

```{r, eval=FALSE}
# don't eval when knitting

criminal_files <- fs::dir_ls("/home/rstudio/courtdata") %>% str_subset("criminal")
print(paste("Processing", length(criminal_files), "files..."))

res <- imap_dfr(criminal_files,  ~ {
  if (.y %% 10 == 0) cat(paste("$", .y, "$"))
  
  suppressWarnings(read_court_file(.x)) %>%
    filter(DispositionCode == "Resolved")
})
cat("\n")

```


```{r}
res$Locality %>% table() %>% sort(decreasing = T) %>% as.data.frame()
```

# Marijuana charges
```{r, eval=FALSE}
# don't eval when knitting

criminal_files <- fs::dir_ls("/home/rstudio/courtdata") %>% str_subset("criminal")
print(paste("Processing", length(criminal_files), "files..."))

res <- imap_dfr(criminal_files,  ~ {
  if (.y %% 10 == 0) cat(paste("$", .y, "$"))
  
  suppressWarnings(read_court_file(.x)) %>%
    filter(stringr::str_detect(CodeSection, "18.2-248.1"))
})
cat("\n")

```

```{r}
res %>%
mutate(year = format(HearingDate, "%Y")) %>%
count(year) %>%
View()
```


