---
title: "Current State of Expungeable Records SCRAPS"
author: "Seth Green"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
# first you need to install these packages
library(httr)
library(tidyverse)
library(data.table)
library(fs)
library(here)
source(here("code", "helper-functions.R"))
```


# Connecting to the data

Here we'll have a function to pull in the data, ideally for a single year. This one works to get a 1000 row sample, but we'll rewrite this once we're on the server to just hit the csv's directly.

```{r}
pull_table_sample <- function(.table) {
  endpoint <- "http://132.145.211.20:8001/criminal-court.csv?"
  my_query <- paste("select * from", .table)
  request <- GET(url = endpoint,
              query = list(
                sql = my_query
              ))
  read_csv(content(request, as = "text", encoding = "UTF-8"))
}

```
And now pull a sample for demo:
```{r}
sample_data <- pull_table_sample("circuit_criminal_2019_anon_00")
print(nrow(sample_data))
head(sample_data)
```

### The full data

The full data is pretty big. Here we filter to the files we care about and then we'll read them one-by-one below.

```{r}
criminal_files <- fs::dir_ls("/home/rstudio/courtdata") %>% str_subset("criminal")
length(criminal_files)
```

Here are two helper functions to pull in only the columns we need. We also rename some of them in the district court files, in order to be consistent with the naming in the circuit court files.


```{r}
.d <- read_court_file(criminal_files[50])
.d
```


# Filter to expungeable cases

This is the real analysis section, we're we filter down to the cases that we care about.

### Defining the cases we care about
This is the most complicated section. Basically, I am defining a list of the "categories" of cases that we care about, as defined by the new law. Each "category" is stored as a query, which is a list with the following elements:

* `reference` -- the section of the new law describing a specific category of case that is expungeable.
* `title` -- a free text sample from that section of the new law, for explanatory purposes.
* Any applicable filters. These elements each contain values that the specified field is allowed to take:
  * `CodeSection` -- The codes for the alleged crime
  * `DispositionCode` -- The relevant dispositions of the cases for the relevant crimes
  * `ChargeType` -- The relevant charge type ("Felony", "Misdemeanor", etc.) if relevant
  
**Note: this may be a partial list.** We should probably go through the new law again with an LAJC lawyer to make sure we have coded this correctly.

```{r echo=FALSE}
QUERIES <- list(
  list(
    # these are "deferred dismissals"
    # these are eligible for automatic _after_ 7 years of no other convictions
    reference = "19.2-392.6 - A",
    title = "Automatic sealing of offenses resulting in a deferred and dismissed disposition.",
    CodeSection = c("4.1-305", "18.2-250.1"),
    DispositionCode = c("Dismissed"),
    HearingPlea = "Guilty"
  ),
  # these are eligible for automatic _after_ 7 years of no other convictions
  list(
    reference = "19.2-392.6 - B",
    title = "Automatic sealing of offenses resulting in a conviction.",
    CodeSection = c("4.1-305", "18.2-96", "18.2-103", "18.2-119", "18.2-120", "18.2-134", "18.2-248.1", "18.2-250.1", "18.2-415"),
    DispositionCode = c("Guilty"),
    ChargeType = c("Misdemeanor") # need to change this to be AmendedCaseType if !is.na(AmendedCaseType)
  ),
  
  # both of these ^ if you're convicted of something else on the same day, then you have to go through petition system
  
  
  
  # this is the petition stuff, complicated caveats below...
  list(
    reference = "19.2-392.12",
    title = "Sealing of offenses resulting in a deferred and dismissed disposition or conviction by petition.",  
    CodeSection = c("18.2-36.1", "18.2-36.2", "18.2-51.4", "18.2-51.5", "18.2-57.2", "18.2-266", "46.2-341.24"),
    DispositionCode = c("Guilty", "Dismissed")
  )
  
  any (Misdemeanor, class 5 felony, or class 6 felony, except * ) AND (felony larceny (18.2-95, maybe others... Rob will get us the others))
    are eligible for expungement ONLY by petition.
    
  note: the 9 misdemeanors mentioned in 19.2-392.6 should _not_ be included here because they are automatic
    
    
    # these can't be expunged at all (because of federal stuff)
    * CodeSection = c("18.2-36.1", "18.2-36.2", "18.2-51.4", "18.2-51.5", "18.2-57.2", "18.2-266", "46.2-341.24"),
)
```


Under the old law, any cases with the following dispositions should be eligible for expungement.

```{r}
# this is 392.8 in new law
# NO 7 year waiting period, this should automatically happen immediately
#   IF you have no prior convictions
#   IF you do have prior convictions, you need to go through the old petition process
OLD_LAW <- c(
      "Dismissed", 
      "Dismissed/Other", 
      "Nolle Prosequi", 
      "Not Guilty/Acquitted", 
      "Not Guilty By Reason Of Insanity", 
      "Not True Bill", 
      "Not Guilty", 
      "Not Guilty/Insanity"
    )


```


7 year rule: "NO convictions of any kind for 7 years from disposition date"

same day rule : if you're convicted of anything that's NOT part of this same category on the same day, then it's no longer automatic and you go to petition under 19.2-392.12

Misdemeanor dismissals (everything under OLD_LAW ^) 19.2-392.11 (also discussed in 19.2-392.8)
 - automatic IF 
   - no convictions on the person's record
   - no arrests or charges in the past 3 years (excluding CaseType == "infraction")
   - AND it's not a "deferral dismissal" (i.e. if plea = "guilty")
 - if NOT ^ then its petition (same as the old law)

Misdemeanor convictions (subject to 7 year rule)
 - 19.2-392.6 - A are automatic (subject to same day rule)
 - 19.2-392.6 - B are automatic (subject to same day rule)
 - 19.2-392.12 are petition (subject to 7 year rule)
   - codes explicitly listed in here are never eligible
   - "provided that such person has (a) never been convicted of a Class 1 or 2 felony or any other felony punishable by imprisonment for life, (b) not been convicted of a Class 3 or 4 felony within the past 20 years, or (c) not been convicted of any other felony within the past 10 years of his petition."

Felonies
 - nothing is automatic
 - 19.2-392.12 are petition (subject to 10 year rule, analogous to 7 year rule)
   - dismissals will be immediately eligible
   - convictions eligible "provided that such person has (a) never been convicted of a Class 1 or 2 felony or any other felony punishable by imprisonment for life, (b) not been convicted of a Class 3 or 4 felony within the past 20 years, or (c) not been convicted of any other felony within the past 10 years of his petition."


Lifetime limit of 2 "sentencing events" that can be sealed. 

There are three different levels we care about:
 - person level
 - the "event" level (rows on the same day)
 - the charge level (this is the row-level)




### Code to filter to those cases

This function takes a single query (i.e. one element from the list we just defined) and filters the input data based on it. We will map over this function below.

```{r}
#' Filter the input data based on a query
#' @param .d A tibble of the criminal court data
#' @param .q A "query list" as defined in the section above.
filter_on_query <- function(.d, .q) {
  .d %>%
    filter(
      CodeSection %in% (.q$CodeSection %||% CodeSection),
      DispositionCode %in% (.q$DispositionCode %||% DispositionCode),
      ChargeType %in% (.q$ChargeType %||% ChargeType)
    ) %>%
    mutate(
      reference = .q$reference,
      title = .q$title
    ) %>%
    select(
      reference,
      title,
      HearingDate,
      person_id,
      CodeSection,
      DispositionCode,
      ChargeType,
      Charge
    )
}
```

Here we map over the queries to get only the relevant cases that fit _any_ of our queries. This is only the sample data. We pull in all of the data below.

```{r}
purrr::map_dfr(QUERIES, function(.query) {
  filter_on_query(sample_data, .query)
}) %>%
  select(
    reference,
    CodeSection,
    DispositionCode,
    ChargeType,
    Charge
  ) %>% knitr::kable()
```

### Conviction-free period

**Very important note: we have not considered the 7-year conviction-free stipulation yet.** This is obviously very important, but was more difficult to query so we are planning to come back to it later. Just be sure to keep that in mind when considering the figures below.


## The full data, summarized by year

This is where we roll it up to what they are likely looking for: by year, how many cases are eligible for expungement and why? For the "and why?" part, we give them the reference section in the new law.

```{r}
years <- as.character(2009:2019)
years_files <- unlist(map(years, ~str_subset(criminal_files, .x)))

# writing out counts as we iterate in purrr
write_lines("file,year,total,old_law,new_law", "counts.csv")
message(paste("Iterating over", length(years_files), "for", length(years), "years..."))

all_df <- purrr::map_dfr(years_files, function(.y) {
  message(paste("Parsing", .y))
  
  if (str_detect(.y, "circuit")) {
    this_df <- read_circuit_file(.y)
  } else {
    this_df <- read_district_file(.y)
  }
  
  # get some counts to write out
  total <- nrow(this_df)
  old_law <- this_df %>%
    filter(DispositionCode %in% OLD_LAW) %>%
    nrow()
  
  # parse the year out of the file name
  this_year <- .y %>% 
    str_replace(".+criminal_", "") %>% 
    str_replace("_anon.+$", "")
  
  # filter to only the rows that match one of our queries
  new_law_df <- purrr::map_dfr(QUERIES, function(.query) {
    filter_on_query(this_df, .query)
  }) %>%
    mutate(
      year = this_year,
      file = .y
    )
  
  # write out the counts
  write_lines(
    paste(.y, this_year, total, old_law, nrow(new_law_df), sep=","),
    "counts.csv",
    append=TRUE
  )
  
  return(new_law_df)
})
nrow(all_df)
```

## Results for the New Law

Total counts of _newly expungeable_ cases by year:
```{r}
all_df %>%
  count(year) %>%knitr::kable()
```

```{r}
all_df %>%
  count(year, reference) %>%
  ggplot(aes(x = year, y = n, fill = reference)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    ylab("Count of cases") +
    ggtitle("Count of Expungeable Cases\n  by year and relevant section of new law")
```

```{r}
all_df %>%
  count(year, CodeSection) %>%
  ggplot(aes(x = year, y = n, fill = CodeSection)) +
    geom_bar(stat = "identity") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    ylab("Count of cases") +
    ggtitle("Count of Expungeable Cases\n  by year and criminal code")

```


## Percentage of total cases that are expungeable (under new law vs old law)

This compares the number of additional cases that will be expungeable under the new law vs. the number of cases that were previously expungeable (_but should now be automatic?_).

```{r}
counts_df <- read_csv("counts.csv", col_types = cols()) %>%
  group_by(year) %>%
  summarize(
    total = sum(total),
    new_law = sum(new_law),
    old_law = sum(old_law)
  ) %>%
  pivot_longer(
    cols = ends_with("law"), 
    names_to = "Eligibility", 
    values_to = "raw_count"
  ) %>%
  mutate(
    Percent = scales::percent(raw_count/total, accuracy = 0.1),
    Count = format(raw_count, big.mark = ",")
  )

```

```{r}
ggplot(counts_df, aes(x = as.factor(year), y = raw_count, fill = Eligibility)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
  xlab("Year") + ylab("Count of cases") +
  ggtitle("Expungeable Cases by year\n  new law vs. old law")
```

The `Percent` column shows the percentage of total cases for that year that fall into each category.
```{r}
counts_df %>%
  select(year, Eligibility, Count, Percent) #%>%knitr::kable(align = "r")
```

